Intentional release of native species undermines ecological stability

### 2022-03-12 rmd/supporting_info.Rmd
edit: manuscript
Squashed commit of the following:
commit 42b7838e55c1587f6a744b0a978c367eac453d43
Author: AkiraTerui <hanabi0111@gmail.com>
Date:   Sat Mar 12 11:50:40 2022 -0500

    edit: affiliation

commit 1933aaca4b7cc7f4d29771b59b45cf512e403b48
Author: aterui <hanabi0111@gmail.com>
Date:   Thu Mar 10 22:06:24 2022 -0500

    edit:  manuscript

commit 0884ad16cc1eb71df364fe1aaf0e7cec0dadd794
Author: aterui <hanabi0111@gmail.com>
Date:   Thu Mar 10 21:32:32 2022 -0500

    edit: manuscript

commit db4e8b6bab9c85a4ba5ba4767ac5442fcf493553
Author: AkiraTerui <hanabi0111@gmail.com>
Date:   T

---
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    includes:
        before_body: title.tex
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
header-includes:
  \pagenumbering{gobble}
  \usepackage{amsmath}
bibliography: reference.bib
csl: science.csl
---

@@

# empirical data

## raw data
source(here::here("code/data_fmt_fishdata.R"))

@@

```{=tex}
\newpage
\pagenumbering{arabic}
```
# Materials and Methods

@@

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012]. In the basic formula without stock enhancement, the population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

```{=tex}
\begin{align}
N_{i,t+1} = N_{i,t}~exp\left[r_i \left(1-\frac{N_{i,t} + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i} \right) \right]exp(\epsilon_{i,t})
\end{align}
```
where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on $i$, $K_i$ the carrying capacity, and $\epsilon_{i,t}$ the species response to stochastic environmental fluctuations that obey an normal distribution $Normal(0, \sigma_{\epsilon}^2)$. We modified this formula to include the effects of stock enhancement (species 1) on reproduction and competition as follows:

```{=tex}
\begin{align}
N_{i,t+1} = [N_{i,t}+\phi_i R_t]exp \left[r_i \left\{1-\frac{N_{i,t}+\alpha_{i1} (N_{1,t} + R_t) + \sum_{j=2}^S \alpha_{ij}N_{j,t}}{K_i} \right\} \right]exp(\epsilon_{i,t})
\end{align}
```
$R_t$ is the number of released individuals, and the parameter $\phi_i$ controls the relative fitness of captive-bred individuals:

@@

```{=tex}
\begin{align}
N_{i,t+1} = [N_{i,t}+\phi_i R_t]exp \left[r_i \left(1-\frac{N_{i,t}+\alpha_{i1}R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i} \right) \right]exp(\epsilon_{i,t})
\end{align}
```
In this model, intrinsic growth rates of unenhanced species $r_{i,i \ne 1}$ and interspecific competition $\alpha_{ij}$ are random draws from a uniform ($r_{i,i \ne 1} \sim Unif(0.5,r_{max})$) and an exponential distribution ($\alpha_{ij} \sim Exp(1/\bar{\alpha})$), respectively. We assumed constant values of intraspecific competition ($\alpha_{ii}=1$), carrying capacity $(K_i = K)$ and the number of releases $(R_t=R)$.

Prior to the main simulation, we performed an extensive sensitivity analysis to identify parameters that strongly influence the relationship between community dynamics and stock enhancement (see **Supplementary text**). We identified five influential parameters (**Table S2-S4**), of which different values were considered in the main simulation as follows: intrinsic growth of enhanced species ($r_1 =$ `r unique(param_set$r1)`), average strength of interspecific competition ($\bar{\alpha}=$ `r unique(param_set$alpha)`), carrying capacity ($K=$ `r unique(param_set$k)`), and relative fitness of captive-bred individuals ($f_R=$ `r unique(param_set$phi)`). Meanwhile, we fixed values of the following parameters: the number of species ($S=$ `r unique(param_set$n_species)`), maximum intrinsic growth rate of unenhanced species ($r_{max}=$ `r unique(param_set$r_max)`), and environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(param_set$sd_env)`). This simulation setup resulted in `r nrow(param_set)` sets of parameter combinations that cover a range of ecological scenarios.

Under each parameter combination, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates). The number of released individuals $R$ was drawn randomly from a uniform distribution for each simulation replicate as $R \sim Unif(0,500)$. We initialized the community with populations of each species drawn from a Poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first `r unique(sim_result$n_warmup)` time steps to allow species to establish populations and reach equilibrium without stock enhancement (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(sim_result$n_burnin)` time steps to reach new equilibrium with selective stock enhancement (burn-in period). We continued the simulation run with stock enhancement and saved the last `r unique(sim_result$n_timestep)` time steps. We obtained the following summary statistics of the whole community $\sum_i^SN_j$: the coefficient of variation (CV), the number of species persist (defied as $N_i > 0.01$ at $t=1600$), the temporal mean ($\mu$), and the SD ($\sigma$). The temporal mean and SD were also calculated for the enhanced species ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately. We summarized values of simulation parameters in **Table S1**.

@@

**Fish stocking.** The stocking of masu salmon began in the 1950s. Hatchery fish are released in spring (fry and smolt stages) and fall (juvenile stage). Although the fish release occurs at multiple locations within a watershed, the stocking information is available only at the watershed level. For each release stage, we assembled annual records of stock enhancement (the number of fish released; 1999-2019) from annual investigations by the Japan Fisheries Research and Education Agency and Salmon and Freshwater Fisheries Research Institute. During the study period, the majority of stocking took place in spring at a fry stage (fry : juvenile : smolt = `r pull(df_stock_stage, ratio)[1]` : `r op(pull(df_stock_stage, ratio)[2])` : `r op(pull(df_stock_stage, ratio)[3])`)

@@

```{=tex}
\begin{align}
N_{s,t} \sim Poisson(\lambda_{s,t} A_{s,t})
\end{align}
```
where $\lambda_{s,t}$ is the expected fish density (individual m^-2^) and $A_{s,t}$ the sampling area (m^2^). Since fish sampling was conducted after the spring stocking of masu salmon, captured fish may include individuals released in the observation year. We explicitly modeled this observation process to avoid biases in estimating temporal community trends:

@@

We used linear regression to quantify the impact of stock enhancement on community dynamics. Although our focus is stock enhancement, each model included climatic and local abiotic variables to account for important environmental differences among sites. Specifically, we developed the following linear regression model taking either the CV, species richness (the number of species present during the observation period), mean, or SD as a response variable $y_s$ with a normal or a Poisson distribution.

@@

```{=tex}
\begin{align}
\gamma_{0,w} &\sim Normal(\mu_{\gamma,w},\sigma_{\gamma}^2)\\
\mu_{\gamma,w} &=\delta_0+\sum_k\delta_k z_{k,w}
\end{align}
```
$\delta_{0}$ is the global intercept and $\delta_k (k > 0)$ are the regression coefficients. The watershed-level predictors $z_{k,w}$ include the yearly stocking of masu salmon (fry + juvenile + smolt; averaged for1999-2019) and ocean productivity (chlorophyll *a* concentration; averaged for 2002-2019). Ocean productivity was included because the majority of the observed species use marine habitats at a certain life stage (i.e., diadromous). The parameter $\sigma_{\gamma}$ accounts for random variation among watersheds that the watershed-level predictors cannot capture. All predictors were standardized (mean = 0, SD = 1) before the analysis.

@@

We performed a sensitivity analysis of the community simulation to identify key simulation parameters that strongly affect the relationships between community dynamics (temporal mean and SD of density) and stock enhancement. We generated 500 sets of parameter combinations by randomly drawing values of seven simulation parameters from uniform distributions (**Table S1**). For each parameter combination, we simulated dynamics of 100 independent communities with varying numbers of stock enhancement $R$ (drawn randomly from $Unif(0, 500)$). This yields a total of $5 \times 10^4$ simulation replicates. In each simulation replicate, we ran 1600 time steps of community dynamics and obtained temporal means and SDs of the whole community ($\sum_i^SN_i$), enhanced species ($N_1$), and unenhanced species ($\sum_{i, i\ne1}^SN_i$) using the last 1000 time steps. The first 600 time steps were discarded as initialization and burn-in periods.

For each parameter combination, we estimated Spearman's rank correlation $\xi$ between the number of releases $R$ and community dynamics, which indicate the effect of stock enhancement under a given ecological context. To examine influences of simulation parameters on $\xi$ (**Table S1**), we developed the following regression model taking $\xi$ as a response variable:

@@

We found that regression coefficients of the following parameters ($\zeta$) never exceeded a value of 0.10: the number of species $S$, maximum intrinsic growth rate of unenhanced species $r_{max}$, environmental stochasticity $\sigma_{\epsilon}$, and relative fitness $f_R$. Therefore, these parameters have little influence on the qualitative relationship between the number of releases and community dynamics. In the main simulation, we fixed values of $S$, $r_{max}$, and $\sigma_{\epsilon}$ (**Table S1**). In the meantime, we varied values of $K$, $r_1$, $f_R$, and $\bar{\alpha}$ as they are potentially influential (**Table S1**). Although $f_R$ had little influence on stocking effects (**Table S1**), we varied this parameter given the significant interest in the relative fitness of captive-bred individuals in the wild.

@@

  table_legend <- paste0("Sensitivity analysis of the community simulation. Parameter estimates of linear regression models (standard errors in parenthesis) are shown. Response variables are Spearman's rank correlation between stock enhancement and either temporal mean density ($\\mu$) or SD ($\\sigma$) of ", c_text[i],". Explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.")

@@

                       "Theoretical predictions for the stocking effect on community dynamics. Rows represent different response variables, and columns show distinct simulation scenarios with different strength of interspecific competition ($\\bar{\\alpha}$) and relative fitness of captive-bred individuals ($f_R$). Other parameters are: ",
                       "number of species $S$ = `r df_param$n_species[",i,"]`; ",
                       "intrinsic growth rate of an enhanced species $r_1$ = `r df_param$r1[",i,"]`; ",
                       "maximum intrinsic growth rate of unenhanced species $r_{max}$ = `r df_param$r_max[",i,"]`; ",
                        "environmental stochasticity $\\sigma_{\\epsilon}$ = `r df_param$sd_env[",i,"]`; ",
                        "carrying capacity $K$ = `r df_param$k[",i,"]`.",
                        "\n\\pagebreak\n")

>>>>

---
output:
  pdf_document:
    latex_engine: xelatex
    toc: true
    includes:
        before_body: title.tex
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
header-includes:
  \pagenumbering{gobble}
  \usepackage{amsmath}
bibliography: reference.bib
csl: nature.csl
---

@@

# empirical data ####

## raw data
source(here::here("code/data_fmt_fishdata.R"))
source(here::here("code/data_fmt_analysis.R"))

@@

```{=tex}
\newpage
\pagenumbering{arabic}
```
# Methods

@@

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012]. In the basic formula without intentional release, the population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

```{=tex}
\begin{align}
N_{i,t+1} = N_{i,t}~exp\left[r_i \left(1-\frac{N_{i,t} + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i} \right) \right]exp(\epsilon_{i,t})
\end{align}
```
where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on $i$, $K_i$ the carrying capacity, and $\epsilon_{i,t}$ the species response to stochastic environmental fluctuations that obey an normal distribution $Normal(0, \sigma_{\epsilon}^2)$. We modified this formula to include the effects of intentional release (species 1) on reproduction and competition as follows:

```{=tex}
\begin{align}
N_{i,t+1} = (N_{i,t}+\phi_i R_t)~exp \left[r_i \left\{1-\frac{N_{i,t}+\alpha_{i1} (N_{1,t} + R_t) + \sum_{j=2}^S \alpha_{ij}N_{j,t}}{K_i} \right\} \right]exp(\epsilon_{i,t})
\end{align}
```
$R_t$ is the number of released individuals, and the parameter $\phi_i$ controls the relative fitness of captive-bred individuals:

@@

```{=tex}
\begin{align}
N_{i,t+1} = (N_{i,t}+\phi_i R_t)~exp \left[r_i \left(1-\frac{N_{i,t}+\alpha_{i1}R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i} \right) \right]exp(\epsilon_{i,t})
\end{align}
```
In this model, intrinsic growth rates of unenhanced species $r_{i,i \ne 1}$ and interspecific competition $\alpha_{ij}$ are random draws from a uniform ($r_{i,i \ne 1} \sim Unif(0.5,r_{max})$) and an exponential distribution ($\alpha_{ij} \sim Exp(1/\bar{\alpha})$), respectively. We assumed constant values of intraspecific competition ($\alpha_{ii}=1$), carrying capacity $(K_i = K)$ and the number of releases $(R_t=R)$.

Prior to the main simulation, we performed an extensive sensitivity analysis to identify parameters that strongly influence the relationship between community dynamics and intentional release (see **Supplementary text**). We identified five influential parameters (**Table S2-S4**), of which different values were considered in the main simulation as follows: intrinsic growth of enhanced species ($r_1 =$ `r unique(param_set$r1)`), average strength of interspecific competition ($\bar{\alpha}=$ `r unique(param_set$alpha)`), carrying capacity ($K=$ `r unique(param_set$k)`), and relative fitness of captive-bred individuals ($f_R=$ `r unique(param_set$phi)`). Meanwhile, we fixed values of the following parameters: the number of species ($S=$ `r unique(param_set$n_species)`), maximum intrinsic growth rate of unenhanced species ($r_{max}=$ `r unique(param_set$r_max)`), and environmental stochasticity ($\sigma_{\epsilon}=$ `r unique(param_set$sd_env)`). This simulation setup resulted in `r nrow(param_set)` sets of parameter combinations that cover a range of ecological scenarios.

Under each parameter combination, we ran `r n_sim` time steps of `r max(sim_result$n_rep)` independent communities (i.e., simulation replicates). The number of released individuals $R$ was drawn randomly from a uniform distribution for each simulation replicate as $R \sim Unif(0,500)$. We initialized the community with populations of each species drawn from a Poisson distribution with a mean of five. We repeated the seeding procedure every 10 time steps over the first `r unique(sim_result$n_warmup)` time steps to allow species to establish populations and reach equilibrium without intentional release (initialization period). After the initialization period, we released $R$ individuals of the enhanced species every time step over the next `r unique(sim_result$n_burnin)` time steps to reach new equilibrium with selective intentional release (burn-in period). We continued the simulation run with intentional release and saved the last `r unique(sim_result$n_timestep)` time steps. We obtained the following summary statistics of the whole community $\sum_i^SN_j$: the coefficient of variation (CV), the number of species persist (defied as $N_i > 0.01$ at $t=1600$), the temporal mean ($\mu$), and the SD ($\sigma$). The temporal mean and SD were also calculated for the enhanced species ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately. We summarized values of simulation parameters in **Table S1**.

@@

**Fish release.** The release of masu salmon began in the 1950s. Hatchery fish are released in spring (fry and smolt stages) and fall (juvenile stage). Although the fish release occurs at multiple locations within a watershed, the release information is available only at the watershed level. For each release stage, we assembled annual records of intentional release (the number of fish released; 1999-2019) from annual investigations by the Japan Fisheries Research and Education Agency and Salmon and Freshwater Fisheries Research Institute. During the study period, the majority of release took place in spring at a fry stage (fry : juvenile : smolt = `r pull(df_stock_stage, ratio)[1]` : `r op(pull(df_stock_stage, ratio)[2])` : `r op(pull(df_stock_stage, ratio)[3])`)

@@

```{=tex}
\begin{align}
N_{s,t} \sim Poisson(\lambda_{s,t} A_{s,t})
\end{align}
```
where $\lambda_{s,t}$ is the expected fish density (individual m^-2^) and $A_{s,t}$ the sampling area (m^2^). Since fish sampling was conducted after the spring release of masu salmon, captured fish may include individuals released in the observation year. We explicitly modeled this observation process to avoid biases in estimating temporal community trends:

@@

We used linear regression to quantify the impact of intentional release on community dynamics. Although our focus is intentional release, each model included climatic and local abiotic variables to account for important environmental differences among sites. Specifically, we developed the following linear regression model taking either the CV, species richness (the number of species present during the observation period), mean, or SD as a response variable $y_s$ with a normal or a Poisson distribution.

@@

```{=tex}
\begin{align}
\gamma_{0,w} &\sim Normal(\mu_{\gamma,w},\sigma_{\gamma}^2)\\
\mu_{\gamma,w} &=\delta_0+\sum_k\delta_k z_{k,w}
\end{align}
```
$\delta_{0}$ is the global intercept and $\delta_k (k > 0)$ are the regression coefficients. The watershed-level predictors $z_{k,w}$ include the yearly release of masu salmon (fry + juvenile + smolt; averaged for1999-2019) and ocean productivity (chlorophyll *a* concentration; averaged for 2002-2019). Ocean productivity was included because the majority of the observed species use marine habitats at a certain life stage (i.e., diadromous). The parameter $\sigma_{\gamma}$ accounts for random variation among watersheds that the watershed-level predictors cannot capture. All predictors were standardized (mean = 0, SD = 1) before the analysis.

@@

We performed a sensitivity analysis of the community simulation to identify key simulation parameters that strongly affect the relationships between community dynamics (temporal mean and SD of density) and intentional release. We generated 500 sets of parameter combinations by randomly drawing values of seven simulation parameters from uniform distributions (**Table S1**). For each parameter combination, we simulated dynamics of 100 independent communities with varying numbers of intentional release $R$ (drawn randomly from $Unif(0, 500)$). This yields a total of $5 \times 10^4$ simulation replicates. In each simulation replicate, we ran 1600 time steps of community dynamics and obtained temporal means and SDs of the whole community ($\sum_i^SN_i$), enhanced species ($N_1$), and unenhanced species ($\sum_{i, i\ne1}^SN_i$) using the last 1000 time steps. The first 600 time steps were discarded as initialization and burn-in periods.

For each parameter combination, we estimated Spearman's rank correlation $\xi$ between the number of releases $R$ and community dynamics, which indicate the effect of intentional release under a given ecological context. To examine influences of simulation parameters on $\xi$ (**Table S1**), we developed the following regression model taking $\xi$ as a response variable:

@@

We found that regression coefficients of the following parameters ($\zeta$) never exceeded a value of 0.10: the number of species $S$, maximum intrinsic growth rate of unenhanced species $r_{max}$, environmental stochasticity $\sigma_{\epsilon}$, and relative fitness $f_R$. Therefore, these parameters have little influence on the qualitative relationship between the number of releases and community dynamics. In the main simulation, we fixed values of $S$, $r_{max}$, and $\sigma_{\epsilon}$ (**Table S1**). In the meantime, we varied values of $K$, $r_1$, $f_R$, and $\bar{\alpha}$ as they are potentially influential (**Table S1**). Although $f_R$ had little influence on release effects (**Table S1**), we varied this parameter given the significant interest in the relative fitness of captive-bred individuals in the wild.

@@

  table_legend <- paste0("Sensitivity analysis of the community simulation. Parameter estimates of linear regression models (standard errors in parenthesis) are shown. Response variables are Spearman's rank correlation between intentional release and either temporal mean density ($\\mu$) or SD ($\\sigma$) of ", c_text[i],". Explanatory variables (i.e., simulation parameters) were standardized (mean = 0, SD = 1) before the analysis.")

@@

                       "Theoretical predictions for the release effect on community dynamics. Rows represent different response variables, and columns show distinct simulation scenarios with different strength of interspecific competition ($\\bar{\\alpha}$) and relative fitness of captive-bred individuals ($f_R$). Other parameters are: ",
                       "number of species $S$ = `r df_param$n_species[",i,"]`; ",
                       "intrinsic growth rate of an enhanced species $r_1$ = `r df_param$r1[",i,"]`; ",
                       "maximum intrinsic growth rate of unenhanced species $r_{max}$ = `r df_param$r_max[",i,"]`; ",
                        "environmental stochasticity $\\sigma_{\\epsilon}$ = `r df_param$sd_env[",i,"]`; ",
                        "carrying capacity $K$ = `r df_param$k[",i,"]`.",
                        "\n\\pagebreak\n")
